---
title: "Biostatistic Assignment 1"
author: "Min Zhu"
date: "`r Sys.Date()`"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(splines)
```

```{r}
# loading data
set.seed(031026) 
cases <- read.table("cases-1.tsv", sep = "\t", header = TRUE )
head(cases)
```

# Task1

```{r}
cases_year_sum <- aggregate(n ~ agegroup + sex, data = cases, FUN = sum)
age_start <- as.numeric(sub("-.*", "", as.character(cases_year_sum$agegroup)))
ord <- order(age_start, cases_year_sum$agegroup)
cases_year_sum <- cases_year_sum[ord, ]
cases_year_sum$agegroup <- factor(as.character(cases_year_sum$agegroup), levels = unique(as.character(cases_year_sum$agegroup)))
head(cases_year_sum,40)
```

```{r}
p1 <- ggplot(cases_year_sum, aes(x = agegroup, y = n, fill = sex)) +
  geom_col(position = position_dodge(width = 0.6), width = 0.6) +
  scale_fill_manual(
  values = c("Male" = "#00BFC4", "Female" = "#F8766D")) +
  labs(x = "Age group", y = "Number of cases", fill = "Sex",
       title = "Number of cases by age group and sex") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(p1)
```

### Observations of graph1: 1.case总数随年龄增加呈现先增多后减少趋势。 2.男女case数整体相近，在老年时女性显著多于男性

# Task2

```{r}
cases_age_sum <- aggregate(n ~ year + sex, data = cases, FUN = sum)
head(cases_age_sum,40)
```

```{r}
p2 <- ggplot(cases_age_sum, aes(x = year, y = n, color = sex, group = sex)) +
  geom_line(size = 1.2) +   
  geom_point(size = 2) +  
  scale_color_manual(values = c("Male" = "#00BFC4", "Female" = "#F8766D")) +
  labs(
    x = "Year",
    y = "Number of cases",
    color = "Sex",
    title = "Number of cases over years by sex"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1))

print(p2)
```

### Observation:case数随年份增长整体呈上升趋势；总体上历年间女性病例数多，男女病例数差距逐渐缩小。

# Task3

```{r}
# loading data
population <- read.table("population-1.tsv", sep = "\t", header = TRUE )
head(population)
population$agegroup <- factor(population$agegroup, levels = unique(population$agegroup))
```

```{r}
p3 <- ggplot(population, aes(x = year, y = agegroup, fill = n_pop)) +
  geom_tile(color = "white") +                   
  scale_fill_gradient(low = "white", high = "red") +
  facet_wrap(~sex) +                             
  labs(
    x = "Year",
    y = "Age group",
    fill = "Population size",
    title = "Population size over age groups and years by sex"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1), 
    strip.text = element_text(face = "bold")          
  )

print(p3)
```

# Task4

```{r}
# Check if all the agegroups are the same
setdiff(unique(cases$agegroup), unique(population$agegroup))  
setdiff(unique(population$agegroup), unique(cases$agegroup))  

# Check if all the years are the same
setdiff(unique(cases$year), unique(population$year))  
setdiff(unique(population$year), unique(cases$year))

# Merge all data
n_merged <- merge(cases, population, by = c("year", "agegroup", "sex"))
names(n_merged)[names(n_merged) == "n"] <- "n_cases"
# order year and agegroup
n_merged$age_start <- as.numeric(sub("-.*", "", n_merged$agegroup))
n_merged <- n_merged[order(n_merged$year, n_merged$age_start), ]
n_merged$age_start <- NULL

head(n_merged, 40)
```

### Answer: Yes, the population file includes the same age groups and calendar years as the file including the number of cases.

```{r}
# sum agegroup first and merge the data
population_age_sum <- aggregate(n_pop ~ year + sex, data = population, FUN = sum)
data_m <- merge(cases_age_sum, population_age_sum, by = c("year", "sex"))
names(data_m)[names(data_m) == "n"] <- "n_cases"
head(data_m,20)
```

# Task5

```{r}
n_merged$IR <- n_merged$n_cases / n_merged$n_pop
data_m$IR <- data_m$n_cases / data_m$n_pop
head(n_merged)
head(data_m)
```

### incidence rate = number of new cases in a period / number of population at risk during the same period

### I think this is the appropriate way of calculating an incidence rate.

# Task6

```{r}
p4 <- ggplot(data_m, aes(x = year, y = IR, color = sex)) +
  geom_point() +                      
  geom_smooth(method = "loess", span = 0.5, se = FALSE) +  
  labs(
    x = "Year",
    y = "Incidence rate",
    color = "Sex",
    title = "Smoothed incidence rate of colon cancer over time by sex"
  ) +
  theme_minimal()

print(p4)
```

### Observation: 女性发病率始终比男性高；随时间逼近，发病率不断增加；男女发病率增长率都经历了先减慢后加快的变化趋势。

# Task 7

```{r}
# Fit Poisson model
poisson_model1 <- glm(n_cases ~ year + sex + offset(log(n_pop)),
                     family = poisson(link = "log"),
                     data = data_m)
summary(poisson_model1)
```

# Task8

```{r}
# IR in 1970
pre_data1 <- data.frame(
  year = 1970,
  sex = c("Male", "Female"),
  n_pop = 1
)
log_n_cases_1 <- predict(poisson_model1, pre_data1, type="link")
IR1 <- exp(log_n_cases_1)
cat("The IR of male in 1970 is",IR1[1],"\n")
cat("The IR of female in 1970 is",IR1[2],"\n")
cat("The difference IR in 1970 between male and female is",abs(IR1[1]-IR1[2]),"\n")

# IR in 2020
pre_data2 <- data.frame(
  year = 2020,
  sex = c("Male", "Female"),
  n_pop = 1
)
log_n_cases_2 <- predict(poisson_model1, pre_data2, type="link")
IR2 <- exp(log_n_cases_2)
cat("The IR of male in 2020 is",IR2[1],"\n")
cat("The IR of female in 2020 is",IR2[2],"\n")
cat("The difference IR in 2020 between male and female is",abs(IR2[1]-IR2[2]))
```

### Assumption: The instance rate grow over calendar years. The difference between male and female also grow.

# Task9

```{r}
##???

```

# Task10

```{r}
dim(model.matrix(~ agegroup*year + sex*year, data=n_merged))
# Crash down
# poisson_model2 <- glm(n_cases ~  agegroup*year + sex*year,
#                       offset = log(n_pop),
#                       family = poisson,
#                       data = n_merged)
# summary(poisson_model2)
```

```{r}
# refer to Task1, just pick up agegroup over 40
n_merged_highr <- subset(n_merged, agegroup %in% c("40-44","45-49","50-54","55-59","60-64","65-69","70-74","75-79","80-84","85-89"))
head(n_merged_highr,40)
# keep group 
dim(model.matrix(~ agegroup*year + sex*year, data=n_merged_highr))
poisson_model2 <- glm(n_cases ~  agegroup*year + sex*year,
                      offset = log(n_pop),
                      family = poisson,
                      data = n_merged_highr)
summary(poisson_model2)
```

```{r}
# forecast
# IR in 1970 in agegroup 70-74 among males and females
pre_data3 <- data.frame(
  year = 1970,
  agegroup = "70-74",
  sex = c("Male", "Female"),
  n_pop = 1
)
# IR in 1970 in agegroup 70-74 among males and females
log_n_cases_3 <- predict(poisson_model2, pre_data3, type="link")
IR3 <- exp(log_n_cases_3)
cat("The IR of male in 1970 in agegroup 70-74 is",IR3[1],"\n")
cat("The IR of female in 1970 in agegroup 70-74 is",IR3[2],"\n")

# IR in 2020 in agegroup 70-74 among males and females
pre_data4 <- data.frame(
  year = 2020,
  agegroup = "70-74",
  sex = c("Male", "Female"),
  n_pop = 1
)

log_n_cases_4 <- predict(poisson_model2, pre_data4, type="link")
IR4 <- exp(log_n_cases_4)
cat("The IR of male in 2020 in agegroup 70-74 is",IR4[1],"\n")
cat("The IR of female in 2020 in agegroup 70-74 is",IR4[2])
```

# Task11

```{r}
# use median represent each agegroup
n_merged_highr$age_mid <- sapply(strsplit(as.character(n_merged_highr$agegroup), "-"),
                                  function(x) mean(as.numeric(x)))
head(n_merged_highr)
dim(model.matrix(~ age_mid*year + sex*year, data=n_merged_highr))
```

```{r}
poisson_model3 <- glm(n_cases ~ age_mid*year + sex*year,
                      family = poisson,
                      offset = log(n_pop),
                      data = n_merged_highr)

summary(poisson_model3)
```

```{r}
# forecast data from splined poisson model
ages <- c(52, 72, 87)
years <- sort(unique(n_merged_highr$year))
sexes <- c("Female","Male")

pred_spl <- expand.grid(
  age_mid = ages,
  year = years,
  sex = sexes
)
pred_spl$n_pop <- 1
pred_spl$fit <- predict(poisson_model3, newdata = pred_spl, type="response")

# observed value
obs_val <- subset(n_merged_highr,
                 agegroup %in% c("50-54","70-74","85-89"))

#plot
p5 <- ggplot() +
      geom_line(data = pred_spl,
                aes(x = year, y = fit,
                    color = sex, linetype = factor(age_mid)),
                size = 1) +
      geom_point(data = obs_val,
                 aes(x = year, y = IR,
                     color = sex, shape = agegroup),
                 size = 2, alpha = 0.7) +
      labs(y = "Incidence rate",
           title = "Predicted and observed incidence rates") +
      theme_bw()

print(p5)

```

# Task12

```{r}
# calculate standard age distribution in 2022
std_pop <- aggregate(
  n_pop ~ sex + agegroup,
  data = subset(n_merged, year == 2022),
  FUN = sum
)
names(std_pop)[names(std_pop) == "n_pop"] <- "std_pop"
head(std_pop)
```

```{r}
mergecases_age_sum <- aggregate(n_cases ~ year + sex, data = n_merged, FUN = sum)
head(mergecases_age_sum)

mergecases_age_sum <- merge(mergecases_age_sum, std_pop,
                 by = c("sex", "agegroup"),
                 all.x = TRUE)
head(mergecases_age_sum,60)
```

\`\`\`
