---
title: "Computer Assignment 2"
author: "Min Zhu & Ziyi Wang"
date: "`r Sys.Date()`"
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exercise 2:2

## 2:2.1 Estimate risk, plot graphs and assess the suitability of the logistic regression model

```{r}
# Create the data frame
logdos <- c(-7.60, -6.22, -4.60, -3.00, -1.39, 0.92)
n <- c(18, 19, 28, 32, 28, 40)
x <- c(1, 2, 4, 9, 12, 32)
#no_tumor <- c(17, 17, 24, 23, 16, 8)


data22 <- data.frame(logdos, x, n)
data22
```

```{r}
# Estimate for each dose separately the risk (probability)
risk <- x / n
risk_data <- data.frame(logdos, risk)
risk_data
```

```{r}
# Plot the risk against log(dos)
plot(risk_data$logdos, risk_data$risk, 
     xlab = "log(dose)", ylab = "Risk (Probability)", 
     main = "Risk vs log(dose)",
     pch = 16, col = "blue", ylim = c(0, 1))
lines(risk_data$logdos, risk_data$risk, col = "blue")

```

```{r}
# Calculate the logodds for tumor
log_odds <- log(risk / (1-risk))
log_odds_data <- data.frame(logdos, log_odds)
log_odds_data
```

```{r}
# Plot the logodds against log(dose).
plot(log_odds_data$logdos, log_odds_data$log_odds, 
     xlab = "log(dose)", ylab = "log(odds)", 
     main = "log(odds) vs log(dose)",
     pch = 16, col = "black")
abline(lm(log_odds ~ logdos, data = log_odds_data), col = "red")

```

From the plot of Risk vs log(dose), it can be seen that the relationship
between logarithmic dose and risk is nonlinear, so directly fitting a
linear model would be inappropriate. However, after calculating the
log-odds, the plot of log(odds) against log(dose) reveals an
approximately linear relationship between the log-odds and the
logarithmic dose. This linearity is a core assumption of the logistic
regression model. Therefore, the logistic regression model is
well-suited for modeling the log-odds in relation to log(dose), as they
demonstrate a linear association.

## 2:2.2 Fit the logistic regression model and interpret the parameter estimates

```{r}
model22 <- glm(x/n ~ logdos, family = binomial(link=logit), weight = n, data = data22)
summary(model22)
```

```{r}
# Extract parameter estimates
intercept_est <- coef(model22)[1]
slope_est <- coef(model22)[2]

cat("Intercept:", round(intercept_est, 4), "\n")
cat("Slope for log(dose):", round(slope_est, 4), "\n")
```

```{r}
# Calculate odds ratio for slope parameter
odds_ratio <- exp(slope_est)
cat("Odds Ratio for log(dose):", round(odds_ratio, 4), "\n")

# Calculate confidence interval for odds ratio
odds_ratio_ci <- exp(confint(model22)[2, ])
cat("95% CI for Odds Ratio:", round(odds_ratio_ci, 4), "\n")
```

$$Logistic\enspace regression:P(Y=1|X=x)=\pi(x)=\frac{exp(\beta_0+\beta_1 x)}{1+exp(\beta_0+\beta_1 x)}$$
Parameter Interpretation: The intercept($\beta_0$) is 0.68702 and the
slope($\beta_1$) is 0.52037 represent the baseline of log odds and the
effect of log dose, respectively. The detailed explanation and proof
shown as follows.

1.  Intercept($\beta_0 = 0.68702$): When log(dose) = 0, the log-odds of
    tumor occurrence is 0.68702, corresponding to a probability of
    $P(Y=1|X=0)=\pi(0)=\frac{exp(\beta_0)}{1+exp(\beta_0)}=\frac{exp(0.68702)}{1+exp(0.68702)}=0.6653$.

2.  Slope parameter($\beta_1 = 0.52037$): For each one-unit increase in
    log(dose), the log-odds of tumor occurrence increases by 0.52037,
    which corresponds to an increase in odds by 68.3%. \begin{equation*}
    \begin{aligned}
    log\enspace OR &=log[\frac{\pi(x+1)}{1-\pi(x+1)}/\frac{\pi(x)}{1-\pi(x)}]\\
    &=logit\enspace \pi(x+1)-logit\enspace \pi(x)\\
    &=\beta_0+\beta_1 (x+1)-(\beta_0+\beta_1 x)\\
    &=\beta_1\\
    &=0.52037
    \end{aligned}
    \end{equation*}

3.  Statistical significance: The slope of the model is significant (p
    \< 0.001), indicating a significant positive effect of dose on tumor
    risk, with the risk increasing as the dose increases.

## 2:2.3 Find covariance matrix, parameter correlation, confidence intervals, and risk at specific dose

```{r}
# Find the covariance matrix for the estimates
cov_matrix <- vcov(model22)
cov_matrix
```

```{r}
# Assess if parameters are correlated
cor_matrix <- cov2cor(cov_matrix)
cor_matrix
```

Correlation between intercept and slope is 0.655691, indicating the
moderate correlation between parameters detected.

```{r}
# Find 95% confidence intervals for the parameters
conf_int <- confint.default(model22,level=0.95)
cat("95% Confidence Intervals:\n")
print(conf_int)
```

95% confidence intervals for intercept is [0.182, 1.192], and 95%
confidence intervals for logdos is [0.354, 0.687].

```{r}
# Find 95% confidence interval for tumor risk at log(dose) = -1.39
# Create prediction data frame
target_logdos <- data.frame(logdos = -1.39)

# Get prediction on log-odds scale with standard error
pred_logit <- predict(model22, newdata = target_logdos, se.fit = TRUE)

```

```{r}
# Extract predicted log-odds and standard error
log_odds_est <- pred_logit$fit
log_odds_se <- pred_logit$se.fit

# Calculate confidence interval on log-odds scale
ci_log_odds_lower <- log_odds_est - 1.96 * log_odds_se
ci_log_odds_upper <- log_odds_est + 1.96 * log_odds_se

# Convert to probability scale using logistic function
risk_est <- plogis(log_odds_est)
ci_risk_lower <- plogis(ci_log_odds_lower)
ci_risk_upper <- plogis(ci_log_odds_upper)

cat("Estimated tumor risk at log(dose)=-1.39 :", round(risk_est, 4), "\n")
cat("95% CI for tumor risk: [", round(ci_risk_lower, 4), ",", round(ci_risk_upper, 4), "]\n")
```

When the log(dose)=-1.39, estimated tumor risk is 0.4909 and the 95%
confidence interval for the risk at that level is [0.3941,0.5885].

## 2:2.4 Perform a Wald-test

$$H_0:\beta_1=0 ; H_a:\beta_1\neq0.$$

```{r}
# The logistic regression model
#model22 <- glm(x/n ~ logdos, family = binomial(link=logit), weight = n, data = data22)
summary_model22 <- summary(model22)

# Extract coefficient estimate and standard error for slope
beta1_hat <- coef(model22)[2]
se_beta1 <- summary_model22$coefficients[2, 2]

cat("Parameter estimates from model:\n")
cat("β₁ (slope) estimate:", round(beta1_hat, 4), "\n")
cat("Standard error of β₁:", round(se_beta1, 4), "\n\n")
```

```{r}
# Manual calculation of Wald test statistic
wald_statistic <- beta1_hat / se_beta1
cat("Manual calculation of Wald statistic:\n")
cat("W =", round(beta1_hat, 4), "/", round(se_beta1, 4), "=", 
    round(wald_statistic, 4), "\n\n")

# Calculate p-value
p_value_manual <- 2 * (1 - pnorm(abs(wald_statistic)))
cat("Manual p-value calculation:\n")
cat("p-value = 2 × (1 - Φ(|W|)) = 2 × (1 - Φ(", round(abs(wald_statistic), 4), "))\n")
cat("Manual p-value:", format.pval(p_value_manual), "\n")

```

```{r}
# Compare with R's output
r_wald_statistic <- summary_model22$coefficients[2, 3]
cat("Wald statistic from R output:", round(r_wald_statistic, 4), "\n")

# Compare with R's p-value
r_p_value <- summary_model22$coefficients[2, 4]
r_p_value
```

We constructed the Wald test, calculated the Wald statistic is
$Z_W=37.457$, and get the p-value \< 0.001. That means by using Wald
test we reject the null hypothesis $\beta_1 =0$, i.e. the log(dose) do
have effect on Tumor risk.

## 2:2.5 Investigate the relationship between sample size and variance of parameter estimates

```{r}
#original covariance matrix
logdos<-c(-7.60,-6.22,-4.60,-3.00,-1.39,0.92)
n<-c(18,19,28,32,28,40)
x<-c(1,2,4,9,12,32)
data22 <- data.frame(logdos,x,n)
model22<-glm(x/n~logdos, family = binomial(link=logit), weights=n, data=data22)
vcov(model22)
for (i in 1:3) {
  data22_new <- data.frame(logdos = data22[,"logdos"], data22[,c(2,3)] * 10^i)
  model22_new<-glm(x/n~logdos, family = binomial(link=logit), weights=n, data=data22_new)
  cat("multiply 10^",i,"\n")
  print(vcov(model22_new))
}
```
