---
title: "Computer assignment 2"
author: "Ziyi Wang & Min Zhu"
date: "`r Sys.Date()`"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
set.seed(031026)
load("proj_data.Rdata")
model1 <- glm(Resultat ~ Alder + Kon + Utbildare, 
              data = data_individ,
              family = "binomial")
summary(model1)
```

```{r, code = readLines("funktioner.R")}
source("funktioner.R")
```

```{r}
y <- matrix(data_individ$Resultat, ncol = 1)
X <- model.matrix(Resultat ~ Alder + Kon + Utbildare, 
                  data = data_individ)
```

## Task 1

```{r}
# Coefficient of glm method
theta_mo <- coef(model1)
theta_mo <- as.numeric(theta_mo)

# Use functions I and NR from Part I
theta0 <- c(0,0,0,0)
max_iter <- 100
for (i in 1:max_iter){
  theta0 <- NR(theta0, niter=1, y, X)
  theta_hat <- theta0
  if (all(round(theta_hat, 2) == round(theta_mo, 2))) {
    cat("niter =", i,"reaches two-digits accuracy")
    break
  }
}
```

```{r}
# Calculate standard error and Wald statistic
fisher_info <- I(theta_hat, y, X)
fisher_inv <- solve(fisher_info)
std_errors <- sqrt(diag(fisher_inv))
wald_stats <- theta_hat/std_errors
print(wald_stats)
```

```{r}
# Compare the z value column of the output
r_z_values <- summary(model1)$coefficients[, "z value"]
#print(r_z_values)

Difference <- round(abs(wald_stats-r_z_values), 6)
print(Difference)
```

### The differences between wald_stats and r_z_values being less than 0.01 confirms that the z-values in R's summary output are indeed Wald statistics.

## Task 2

```{r}
# Under H0
eta <- NR(theta0 = c(0,0,0),niter=10, y=y, X=X[,-3])
eta
L_eta <- L(eta,y,X=X[,-3])

# Without H0 restrction 
L_raw <- L(theta_hat,y,X)

# Calculate generalized likelihood ratio
ratio <- L_raw/L_eta
W <- 2*log(ratio)
cat("generalized likelihood ratio statistics:", W, "\n")

# p-value
p_value <- pchisq(W, df = 1, lower.tail = FALSE)
cat("P-values:", p_value, "\n")
```
### Since the Wald statistic follows an asymptotic standard normal distribution, its square follows a chi-square distribution with one degree of freedom. The generalized likelihood ratio statistic also asymptotically follows a chi-square distribution, and in this problem its parameter (q) is one. Therefore, both statistics are of the same order of magnitude.

## Task 3

```{r}
# Fit the model under H0:θ=(θ_intercept,θ_Kon)=(0,0)
X_H0 <- X[, c(2, 4)]

# Estimate η under H0
eta_H0 <- NR(theta = c(0, 0), niter = 10, y = y, X = X_H0)

# Construct full parameter vector
theta0 <- c(0, 0)  # (intercept, Kon) = (0, 0)
beta_H0 <- c(theta0[1], eta_H0[1], theta0[2], eta_H0[2])

# Compute the score vector
score_H0 <- S(beta_H0, y, X)

# Compute Fisher information matrix
info_H0 <- I(beta_H0, y, X)

# Compute the generalized score statistic
# T_S = S(θ0, η_ML)^T * I(θ0, η_ML)^{-1} * S(θ0, η_ML)
T_S <- t(score_H0) %*% solve(info_H0) %*% score_H0
T_S <- as.numeric(T_S)
cat("Generalized Score Statistic T_S:", T_S, "\n")
```
```{r}
# Compute p-value (asymptotically χ²(2))
p_value <- 1 - pchisq(T_S, df = 2)
cat("P-value for H0:", p_value, "\n")
```
## Task 4

```{r}
# Compute wald interval
# Z-crit
z_crit <- qnorm(0.975)

# Wald interval
Wald_L <- theta_hat[3] - z_crit * std_errors[3] # Lower Confidence Limit
Wald_U <- theta_hat[3] + z_crit * std_errors[3] # Upper Confidence Limit
cat("Wald interval of theta_kon is:", Wald_L, "to", Wald_U)
```
```{r}
# # Compute profile likelihood at a fixed theta_Kon=t:
# function 
Profile_likelihood <- function(t){
  theta.Kon <- t
  profile <- glm.fit(x= X[, -3], y = y,
                     offset = theta.Kon * X[,3],
                     family = binomial())
  pro_coeff<- profile$coeff
  # plug back
  linear_pred <- profile$linear.predictors
  probs <- binomial()$linkinv(linear_pred)
  log_lik <- sum(dbinom(y, size = 1, prob = probs, log = TRUE))
  pro_lik <- exp(log_lik)
  return (pro_lik)
}

# test
LP_thetakon <- Profile_likelihood(0.5)
LP_thetakon
```

```{r}
# Compute Estimated likelihood
eta_global <- coef(model1)[-3]
Estimated_likelihood <- function(t){
  theta.Kon <- t
  linear_pred <- X[,-3] %*% eta_global + X[,3] * theta.Kon
  probs <- binomial()$linkinv(linear_pred)
  log_lik <- sum(dbinom(y, size = 1, prob = probs, log = TRUE))
  pro_lik <- exp(log_lik)
  return (pro_lik)
}

# test
Le_thetakon <- Estimated_likelihood(0.5)
Le_thetakon
```
```{r}
# Compute Profile and Estimated value
theta_range <- seq(-0.5,0.5,by=0.001)
LP_thetakon <- sapply(theta_range, Profile_likelihood)
Le_thetakon <- sapply(theta_range, Estimated_likelihood)
```

```{r}
# Find 95% interval
LP_max <- max(LP_thetakon)
Pro_relative <- LP_thetakon / LP_max
Re_p <- qchisq(0.95,df=1)
threshold <- exp(-Re_p /2)
LP_CI95 <- threshold * LP_max
LP_CI95
```

```{r}
## Draw
# y_range
y_lim_all <- range(c(LP_thetakon, Le_thetakon))

# Profile likelihood
plot(theta_range, LP_thetakon,
     type = "l", 
     lwd = 3,                       
     col = "blue", 
     xlab = expression(theta[Kon]),      
     ylab = "Likelihood Value",  
     main = "Profile vs Estimated Likelihood Function",
     ylim = c(min(y_lim_all), max(y_lim_all)),
     las = 1                        
)

# Estimated likelihood
lines(theta_range, Le_thetakon, 
      lwd = 3, 
      col = "red", 
      lty = 2)

# 95% confidence interval
abline(h = LP_CI95, 
       col = "darkgreen", 
       lty = 3,           
       lwd = 2)

# Wald interval
abline(v = c(Wald_L, Wald_U), 
       col = "orange", 
       lty = 3,  
       lwd = 2)

# Background & Label
grid(col = "gray", lty = 1, lwd = 0.5)
legend("topright",                 
       legend = c("Profile", 
                  "Estimated", 
                  "95% thereshold",
                  "Wald interval"),
       col = c("blue", "red", "darkgreen","orange"),
       lwd = c(3, 3, 2, 2),
       lty = c(1, 2, 3, 3),
       bty = "n"                    
)

```

